---
title: "MD5"
author: "Wayne Frederick II"
format: html
embed-resources: true
editor: visual
---

## Question 1

```{r}
#| message: false
library(tidyverse)

exercise_data <- read_csv("https://raw.githubusercontent.com/vaiseys/dav-course/main/Data/visualize_data.csv")
glimpse(exercise_data)
```

If people exercise more than they should have a lower BMI.

```{r}
cor(exercise_data$Exercise, exercise_data$BMI)
```

Shows the negative correlation between exercise and bmi.

```{r}
ggplot(data = exercise_data, mapping = aes(x = Exercise, y = BMI)) + geom_point()
```

## Question 2

```{r}
#| message: false
library(causact)
glimpse(corruptDF)
```

```{r}
?corruptDF
```

CPI2017: A country's CPI score shows what the country thinks about the level of public sector corruption on a scale of 0-100.

HDI2017: HDI score show how well human development is going across countries

## Question 3

```{r}
ggplot(data = corruptDF, mapping = aes(x = CPI2017, y = HDI2017)) + geom_point()
```

There is a positive correlation between CPI and HDI in 2017. As one increases the other does as well on average

## Question 4

```{r}
ggplot(data = corruptDF, mapping = aes(x = HDI2017, y = CPI2017)) + geom_point() + geom_smooth(method = "lm")
```

I like lm better because it is a more straight forward and easier to read graph.

## Question 5

```{r}
ggplot(data = corruptDF, mapping = aes(x = HDI2017, y = CPI2017, fill = region)) + geom_point() + geom_smooth(method = "lm") + facet_wrap(~region)
```

It was difficult analysing the data without facet wrapping it because all of it was layered. used fill to add the color and geom smooth to collect the data into one. Facet wrap allowed me to separate the data more clearly into individual graphs.

## Question 6

```{r}
ggplot(data = corruptDF, mapping = aes(x = HDI2017, y = CPI2017, fill = region)) + geom_point() + geom_smooth(method = "lm") + facet_wrap(~region) + scale_x_reverse()
```

## Question 7

```{r}
ggplot(data = corruptDF, mapping = aes(x = HDI2017, y = CPI2017, fill = region)) + geom_point() + geom_smooth(method = "lm") + facet_wrap(~region) + scale_x_reverse() + labs(title = "Relationship betweend HDI and CPI", subtitle = "Data points are countries", caption = "Source: corruptDF")
```

## Question 8

```{r}
ggplot(data = corruptDF, mapping = aes(x = HDI2017, y = CPI2017, fill = region)) + geom_point() + geom_smooth(method = "lm") + facet_wrap(~region) + scale_x_reverse() + labs(title = "Relationship betweend HDI and CPI", subtitle = "Data points are countries", caption = "Source: corruptDF")

ggsave("HDI-CDI.pdf")
```
