---
title: "DV5"
author: "Wayne Frederick II"
format: html
embed-resources: true
editor: visual
---

## 

```{r}
library(tidyverse)
# Read in the data 
wncaa <- read_csv("https://raw.githubusercontent.com/vaiseys/dav-course/main/Data/wncaa.csv")

# Glimpse the data 
glimpse(wncaa)
view(wncaa)
```

## Question 1

```{r}
champs <- wncaa %>%
  filter(tourney_finish == "Champ")
```

```{r}
percentwins <- champs %>% 
  group_by(school) %>% 
  summarise(N = n()) %>% 
  mutate(percent = N / sum(N),  pct = round((percent*100), 0))
  
percentwins
```

```{r}
ggplot(percentwins, 
       mapping = aes(x = school, y = pct)) + 
  geom_col(position = "dodge2", fill = "purple") +
  scale_y_continuous(limits = c(0, 100)) +
  coord_flip() + labs(title = "Percentage of Tournaments won by school", x = "School", y = "Tournament" )
```

UConn and Tennessee have the highest percentage of wins.

## Question 2

```{r}
thechamps <- unique(champs$school)

winners <- wncaa %>% 
  filter(school %in% thechamps)

ggplot(winners, mapping = aes(x = school, y = seed,  fill = school)) + geom_boxplot(outlier.shape = NA) + geom_jitter(alpha = 0.2) + coord_flip() + labs(title = "Distribution of Seeds by School", x = "School", y = "Seed", caption = "source: WNCAA")
```

The boxplot shows the distribution of seeds by schools that have won a championship. Now to make it easier to interpret, I am reordering the data from lowest average seeds to highest average seeds.

```{r}
winners <- winners |>   
  group_by(school) %>%  
  mutate(mean_seed = mean(seed)) %>% 
  ungroup()

ggplot(winners, 
       aes(x = reorder(school, -mean_seed), y = seed, fill = school)) + geom_boxplot(outlier.shape = NA) + geom_jitter(alpha = 0.3) + coord_flip() + labs(title = "Distribution of Seeds by School", x = "School", y = "Seed", caption = "source: WNCAA")
```

Tennessee and UConn were averaging the 1 seed so it makes sense that they have the best tournament win percentage.

I am now using geom_violin() to visualize the data.

```{r}
ggplot(winners, 
       aes(x = reorder(school, -mean_seed), y = seed, fill = school)) +
  geom_violin() +
  coord_flip() +
  labs(title = "Distribution of Seeds by School", x = "School", y = "Seed", caption = "source: WNCAA")
```

I think the graph with the boxplots tells more because the distribution of data is more clear. The violin graph is hard to read because a lot of them are in the same range so they are hard to tell apart.

## Question 3 

```{r}
ggplot(winners, 
       aes(x = reorder(school, -mean_seed), y = seed, fill = school)) +
  geom_point() +
  geom_jitter(alpha = 0.5) +
  coord_flip() +
  labs(title = "Distribution of Seeds by School", x = "School", y = "Seed", caption = "source: WNCAA")
```

The scatterplot does not show the medians, ranges outlies, or distributions well.

## Question 4

```{r}
winners_mean_sd <- winners |> 
  group_by(school) |> 
  summarize(across(where(is.numeric),
                   list(mean = mean,
                        sd = sd)))
```

```{r}
ggplot(winners_mean_sd,
        mapping = aes(x = reorder(school, reg_percent_mean, na.rm = TRUE),
                      y = reg_percent_mean)) +
  geom_point() +
  coord_flip() +
  labs(title = "Regular Season Win Percentage by School", x = "School", y = "Average Regular Season Win Percentage", caption = "source: WNCAA")
```

THis graph shows that the highest average regular season win percentage was Uconn at 91.54%. THe lowest was Texas A&M with 73.99%. Most of the data points are around 77-80%.

```{r}
ggplot(winners_mean_sd,
        mapping = aes(x = reorder(school, reg_percent_mean, na.rm = TRUE),
                      y = reg_percent_mean)) +
  geom_pointrange(mapping = aes(ymin = reg_percent_mean - reg_percent_sd,
       ymax = reg_percent_mean + reg_percent_sd)) +
  coord_flip() +
  labs(title = "Regular Season Win Percentage by School", x = "School", y = "Average Regular Season Win Percentage", caption = "source: WNCAA")

```

This shows the standard deviation. The school with the. lowest one is Texas A&M. They have the least variation. The schools hover around 67-80%.

```{r}
ggplot(winners_mean_sd,
        mapping = aes(x = reorder(school, reg_percent_mean, na.rm = TRUE),
                      y = reg_percent_mean)) +
  geom_linerange(mapping = aes(ymin = reg_percent_mean - reg_percent_sd,
       ymax = reg_percent_mean + reg_percent_sd)) +
  coord_flip() +
  labs(title = "Regular Season Win Percentage by School", x = "School", y = "Average Regular Season Win Percentage", caption = "source: WNCAA")
```

This graph also shows the standard deviation. The mean is not visible on this graph like it is on the previous one.
