---
title: "MD06X"
author: "Wayne Frederick II"
format: html
embed-resources: true
editor: visual
---

```{r}
library(tidyverse)
library(moderndive)
theme_set(theme_minimal())
```

```{r}
data(bikes, package = "bayesrules")
glimpse(bikes)
```

## Question 0

There are 500 rows in the dataset. Each row is the ridership data for one day. The dataset ranges from January 2011-December 2012. The highest number of rides in a day is 6946. The highest wind speed is 34 mph.

## Question 1

```{r}
get_correlation(rides ~ temp_feel, data = bikes) 
  round(2)
```

The correlation between \# of rides and the temperature is 0.58. When temp_feel increases by 1 standard deviation then rides will increase by 0.58 of a standard deviation.

```{r}
get_correlation(rides ~ windspeed,
                data = bikes) %>% round(2)
```

The correlation between \# of rides and the windspeed is -0.19. When windspeed increases by 1 standard deviation the rides will decrease by 0.19 standard deviation.

## Question 2

```{r}
bikes <- bikes %>% 
  mutate(wind_kph = windspeed * 1.61)
```

```{r}
get_correlation(windspeed ~ wind_kph,
                data = bikes)
```

The correlation between windspeed in kph and mph is 1 because there is a unit conversion that you have to take into account. Miles are 1.61x kilometers.

## Question 3

```{r}
mphmodel <- lm(rides ~ windspeed,
                data = bikes)
get_regression_table(mphmodel) %>% 
  select(term, estimate)
```

```{r}
kphmodel <- lm(rides ~ wind_kph,
                data = bikes)
get_regression_table(kphmodel) %>% 
  select(term, estimate)
```

Both intercepts are the same at 4205.065. The coefficients are off by 1.61x.

```{r}
-55.523 / -34.48600
```

## Question 4

```{r}
 4205.065-34.486*20
```

3515.345 is the predicted ridership for a 20mph wind day.

```{r}
4205.065 - 55.523*20
```

3094.605 is the predicted amount of riders on a 20kph wind day.

## Question 5

```{r}
bikes <- bikes %>% 
  mutate(temp_c = (temp_feel-32)*(5/9))
```

```{r}
mult_model <- lm(rides ~ wind_kph + temp_c,
                 data = bikes)
get_regression_table(mult_model) %>% 
  select(term, estimate)
```

Every 1°C increase in temperature there is an increase of 143 in ridership. For every 1kph increase in wind speed,there is a decrease of 20 in ridership.

The intercept shows that ridership on a 0mph wind day with a temperature of 0°C would be 942.

## Question 6

```{r}
pred_df <- tibble(
  situation = 1:3,
  temp_c = c(25, 15, 10),
  wind_kph = c(15, 5, 40)
)
```

For scenario 1 it is 4226 for scenario 2 it is 2992 and for scenario 3 it is 1581

## Question 7

```{r}
bikes <- bikes %>% mutate(weekend = as.factor(weekend))
weekendmodel <- lm(rides ~ wind_kph + weekend + temp_c,
                    data = bikes)
get_regression_table(weekendmodel) %>%
  select(term, estimate)
```

The intercept coefficients are used in the same way. It is expected that ridership will be 714 fewer than on a weekday.

## Question 8

```{r}
library(ggeffects)
predictions <- ggpredict(weekendmodel,
                         terms = c("weekend"),
                         ci_level = NA)
predictions
```

When looking at the temp and wind speed average the expected ridership a weekdays is 3683 riders and for a weekend it is 2970.

## Question 9

```{r}
bikerides <- predict(weekendmodel)
residuals <- bikes$rides - bikerides
broom::tidy(residuals)
residuals
```

A large residual on day 439. Had an absolute value of 3940.346. On that had hurricane sandy was about to hit in that area.
