---
title: "MD6"
author: "Wayne Frederick II"
format: html
embed-resources: true
editor: visual
---

```{r}
library(broom)
library(tidyverse)
library(moderndive)
library(dplyr)
library(patchwork)
library(ggthemes)
# Set our ggplot theme from the outset
theme_set(theme_light())
# Read in the data 
gender_employment <- read_csv("https://raw.githubusercontent.com/vaiseys/dav-course/main/Data/gender_employment.csv")

# Glimpse at the data 
glimpse(gender_employment)
```

```{r}
gender_employment%>% 
  ggplot(aes(x = year, y = wage_percent_of_male)) +
  geom_jitter(alpha = 0.1) + 
  geom_smooth(method = "lm") + 
  labs(title = "Women's earnings with respect to men's", 
       y = "% of Men's Income", 
       x = "Year")
```

## Question 1

```{r}
gender_employment <- gender_employment %>% 
  mutate(major_category = as.factor(major_category), 
         major_category = relevel(major_category, ref = "Management, Business, and Financial"))
```

```{r}
parallel_model <- lm(wage_percent_of_male ~ year + major_category,
                     data = gender_employment)
broom::tidy(parallel_model)
```

```{r}
-306.7184+ 0.1922*2015+3.3261  
```

Sales and Office occupations had 83.9% of male income in 2015.

```{r}
-306.7184+ 0.1922*2016+ 5.7563  
```

Service occupations had 86.5% of male income in 2015.

## Question 2

```{r}
gender_employment %>% 
  ggplot(aes(x = year, y = wage_percent_of_male)) +
  geom_jitter(alpha = 0.1) + 
  geom_smooth(method = "lm") + facet_wrap(~major_category)
  labs(title = "Women's earnings with respect to men's", 
       y = "% of Men's Income", 
       x = "Year")
```

Their is a lot of difference in the wage percent of male across the categories so we cannot use the parallel trends assumption.

## Question 3

```{r}
interactionmodel <- lm(wage_percent_of_male ~ year*major_category, data = gender_employment) 

broom::tidy(interactionmodel)
```

```{r}
-1370.4719187 + 0.7202561*2016 + 1002.8532317*1 - 0.4946783*2016*1
```

Our model shows that the Computer, Engineering, and Science jobs made 87.14% of male income in 2016.

```{r}
-1370.4719187 + 0.7202561*2016 + 2137.6501474*1 - 1.0581075*2016*1
```

The model estimates that service jobs made 86.07% of male income. The main difference is that there are more components needed to make this calculation compared to the previous model

## Question 4

We would choose to use a model that assumes parallel trends because it can be useful to have a baseline for the whole data set which will show the relationship well.

## Question 5

```{r}
simple_fit <- lm(wage_percent_of_male ~ year,
                 data = gender_employment)
broom::tidy(simple_fit)
```

For every year the wage percent of male counterparts goes up by 0.2%.

```{r}
gender_employment %>% 
  select(year, wage_percent_of_male, percent_female) %>% 
  cor(use = "complete.obs")
```

This explains relationship between two variables that we want to examine. As the year increases the wage percent of males increases by 0.24% and the percent of workforce that is female increases by 0.005%.

```{r}
multiple_fit <- lm(wage_percent_of_male ~ year*percent_female,
                   data = gender_employment)
broom::tidy(multiple_fit)
multiple_fit
```

The proportion of women who work in occupation is increasing by 10%. The relationship between year and percent_female is less than 1%.

## Question 6

```{r}
simple_glanced <- broom::glance(simple_fit)
simple_glanced$r.squared
```

r\^2 explains tells us how much the variation of a dependent variable is explained by the independent variable.

```{r}
multiple_glanced <- broom::glance(multiple_fit)
multiple_glanced$r.squared
```

r\^2 is larger in this case so that means more of the variation of the dependent variable is relying on the independent variable hence the larger number.
