---
title: "MD5"
author: "Wayne Frederick II"
format: html
embed-resources: true
editor: visual
---

```{r}
library(tidyverse)

exercise_data <- read_csv("https://raw.githubusercontent.com/vaiseys/dav-course/main/Data/visualize_data.csv")
glimpse(exercise_data)
```

If people exercise more than they should have a lower BMI.

```{r}
cor(exercise_data$Exercise, exercise_data$BMI)
```

Shows the negative correlation between exercise and bmi

```{r}
ggplot(data = exercise_data, mapping = aes(x = Exercise, y = BMI)) + geom_point()
```

```{r}
library(causact)
glimpse(corruptDF)
```

```{r}
?corruptDF
```

CPI2017: A country's CPI score shows what the country thinks about the level of public sector corruption on a scale of 0-100.

HDI2017: HDI score show how well human development is going across countries

## Question 3

```{r}
ggplot(data = corruptDF, mapping = aes(x = CPI2017, y = HDI2017)) + geom_point()
```

There is a positive correlation between CPI and HDI in 2017. As one increases the other does as well on average

## Question 4

```{r}
ggplot(data = corruptDF, mapping = aes(x = HDI2017, y = CPI2017)) + geom_point() + geom_smooth(method = "lm")
```

## Question 5

```{r}
ggplot(data = corruptDF, mapping = aes(x = HDI2017, y = CPI2017, fill = region)) + geom_point() + geom_smooth(method = "lm") + facet_wrap(~region)
```

I used fill to add the color and geom smooth to collect the data into one. Facet wrap allowed me to separate the data more clearly into individual graphs

## Question 6

```{r}
ggplot(data = corruptDF, mapping = aes(x = HDI2017, y = CPI2017, fill = region)) + geom_point() + geom_smooth(method = "lm") + facet_wrap(~region) + scale_x_reverse()
```

## Question 7

```{r}
ggplot(data = corruptDF, mapping = aes(x = HDI2017, y = CPI2017, fill = region)) + geom_point() + geom_smooth(method = "lm") + facet_wrap(~region) + scale_x_reverse() + labs(title = "Relationship betweend HDI and CPI", subtitle = "Data points are countries", caption = "Source: corruptDF")
```

## Question 8

```{r}
ggplot(data = corruptDF, mapping = aes(x = HDI2017, y = CPI2017, fill = region)) + geom_point() + geom_smooth(method = "lm") + facet_wrap(~region) + scale_x_reverse() + labs(title = "Relationship betweend HDI and CPI", subtitle = "Data points are countries", caption = "Source: corruptDF")

ggsave("HDI-CDI.pdf")
```
